---
title: "Plotting different figures about production linkages"
author: "Jose Ignacio Gonzalez Rojas"
date: "30/05/2020"
output:
  html_document:
    df_print: paged
    number_sections: yes    
    toc: yes
  pdf_document:
    df_print: paged
    number_sections: yes
    toc: yes
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{mathpazo}
---

# General information

The purpose of this code is to create the necessary variables in order to plot a histogram with the count of the sellers per buyer in a specific year, also, this is calculated by sector. These to histograms are also plot with the weights provided by the variable 'buyer_sales'. Also, the share of sellers per buyer and buyers per sellers that he/she still buys/sells from/to between two years. The first one is also weighted by the variable 'buyer_sales'. Finally, the latter variables are plotted in a scatter plot with polynomial fits included. 

## Input files

1. `0-Raw_Data/Sales/sales_new.csv`

## Output files

None. The graphics are kept in code and can be exported to whichever format it's wanted, as well as the dataframes. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
options(tinytex.verbose = TRUE)
```


```{r packages, message=FALSE, warning=FALSE, cache=FALSE, include=FALSE, results='hide'}
# libraries
# vector of libraries to be used

libs <- c(
  "dplyr", "tidyr", "styler", "lintr", "readr", "janitor", "ggplot2",
  "broom", "geosphere", "reshape", "scales", "lubridate", "plotrix"
)

# Installing libraries in case they are not installed
new_packages <- libs[!(libs %in% installed.packages()[, "Package"])]
if (length(new_packages)) install.packages(new_packages)

# Load corresponding libraries

lapply(libs, require, character.only = TRUE)
```

# Importing data to R

In this step, the data needed it's imported into R. Note that the data sets are
in Comma-Separated Values format, so that's why the library "readr" was loaded. The year variable is loaded as DateTime to allow *ggplot2* to detect it as a date. This will make easier the labelling of the x-axis. 

```{r data-importing, echo=TRUE}
# Main inputs

sales_new <- read_csv("0-Raw_Data/Sales/sales_new.csv")
```

## Cleaning the data

The library "janitor" helps standardizing the variable names to the
*snake_case*'s format.

```{r cleaning-variable-names, echo=TRUE}
sales_new <- janitor::clean_names(sales_new)
```

# Number of sellers

## Data processing and statistics

First, the number of sellers per buyer is calculated, this will be helpful to create the plots in adition to the 'buyer_sales' variable as weights. Then, the number of sellers per buyer in each sector is computed. This will be done to a subset of the data corresponding to the year 2001. 

```{r data-manipulation}
yr <- 2001
# note1: better if comments are also in the code
# note2: better if variable names are informative

# number of sellers per buyer
sellers_per_buyer <- sales_new %>%
  filter(year == yr) %>%
  select(buyer, buyer_sales, year) %>%
  group_by(buyer, year) %>%
  mutate(n_sellers = n()) %>%
  unique() %>%
  ungroup()

# number of seller sectors per buyer
sectors_per_buyer <- sales_new %>%
  filter(year == yr) %>%
  select(seller_sector, buyer, buyer_sales, year) %>%
  group_by(buyer, seller_sector, year) %>%
  mutate(n_seller_sectors = n()) %>%
  ungroup() %>%
  select(buyer, buyer_sales, year, n_seller_sectors) %>%
  group_by(buyer, year) %>%
  unique() %>%
  ungroup()

# weighted number of sellers per buyer (?)
# note: each buyer appears as many times as the number
# of sellers already. I think there is double-counting
# in the following formula
sales_new2 <- sales_new %>%
  filter(year == yr) %>%
  select(buyer, seller, buyer_sales) %>%
  count(buyer, wt = buyer_sales)
```

## Histograms

These histograms correspond to the task 3.1. The first two graphics show the frequency of how many sellers per buyer are there, the first is unweighted and the second one is weighted by the variable "buyer_sales". 

```{r histograms}

##
## 1. Number of sellers per buyer
##

## unweighted
# this is another way
h <- hist(sellers_per_buyer$n_sellers, 
          main = "Unweighted number of sellers", 
          freq = FALSE
          )

## weighted
h <- weighted.hist(sellers_per_buyer$n_sellers,
  sellers_per_buyer$buyer_sales,
  freq = FALSE,
  main = "Weighted number of sellers",
  breaks = seq(
    0, max(sellers_per_buyer$n_sellers), max(sellers_per_buyer$n_sellers) / 10
  )
)

graph1 <- ggplot(data = sellers_per_buyer, aes(x = n_sellers)) +
  geom_histogram(bins = 10, boundary = 0) +
  scale_x_continuous(breaks = seq(0, max(sellers_per_buyer$n_sellers), max(sellers_per_buyer$n_sellers) / 10))
graph1

##
## 2. Number of sectors per buyer
##

## unweighted
h <- hist(sectors_per_buyer$n_seller_sectors,
  main = "Unweighted number of sellers",
  breaks = seq(
    0, max(sectors_per_buyer$n_seller_sectors) + 1, 
    max(sectors_per_buyer$n_seller_sectors) / 10
  ),
  freq = FALSE
)

## weighted
h <- weighted.hist(
  sectors_per_buyer$n_seller_sectors,
  sectors_per_buyer$buyer_sales,
  freq = FALSE,
  main = "Weighted number of sellers",
  breaks = seq(
    0, max(sectors_per_buyer$n_seller_sectors) + 1, 
    max(sectors_per_buyer$n_seller_sectors) / 10
  )
)
```

# persistence of the productive linkages between two year: buyers and sellers. 

## persistance between two years for sellers and buyers. 

This code will work for any time series.  

```{r persistence_producitive}
# all the unique years from the 'year' variable are put into a vector
years <- unique(sales_new$year)
# the last year is popped because the for-loop needs to stop in the year $T-1$ to calculate the difference between firms in the year $T$.
years <- years[-length(years)]

# two empty dataframes are created to store the averages of the share of firms that remain from one year to the next one
persistence_buyers <- read.csv(text = "year, mean_share, mean_share_wt")
persistence_sellers <- read.csv(text = "year, mean_share")

##
## 1. BUYERS
##

# variables will be created for every year $t-1$ and the next one $t$ for the buyers and sellers
for (i in years) {
  # for each year $t-1$ the set of links between firms is assigned to a dataframe with their respective weights per buyer corresponding to the buyer_sales in year $t-1$. Then, the number of sellers per buyer is calcultated.
  temp_df1 <- sales_new %>%
    filter(year == i) %>%
    select(buyer, buyer_sales) %>%
    add_count(buyer) %>%
    distinct()

  # the firms of interest are the ones that trade in year $t-1$ and year $t$, we're not interested in the firms that leave the market in year $t$. So, the list of the buyers in year $t-1$ and year $t$ are attached to their respective vectors.
  temp_df2 <- sales_new %>%
    filter(year == i) %>%
    select(buyer) %>%
    distinct()

  temp_df3 <- sales_new %>%
    filter(year == i + 1) %>%
    select(buyer) %>%
    distinct()

  # then, the buyers that trade in both years are kept
  temp_df2 <- inner_join(temp_df2, temp_df3, by="buyer")

  # from the original list of buyers with their respective weights, the only buyers that are kept are the ones which trade in both years, in order to discard the ones that go out of business. With that in mind, the persistence between two years will be measured.
  temp_df1 <- left_join(temp_df2, temp_df1, by="buyer")

  # Another set of the sellers and buyers in the year $t-1$ and the year $t$ is created.
  temp_df4 <- sales_new %>%
    filter(year == i) %>%
    select(buyer, seller)

  temp_df5 <- sales_new %>%
    filter(year == i + 1) %>%
    select(buyer, seller)

  # only the links between firms that are present in the two consecutive years are kept
  temp_df4 <- inner_join(temp_df4, temp_df5, by = c("buyer", "seller")) %>%
    distinct() %>%
    count(buyer)
  # the number of sellers per buyer is computed in the last list

  # the number of sellers per buyer that remained trading with eachother in year $t$ is matched with the number of sellers per buyer in year $t-1$. The missing values are changed by 0's
  temp_df1 <- left_join(temp_df1, temp_df4, by = "buyer")

  temp_df1[is.na(temp_df1)] <- 0

  # the observations are matched and then the share of firms that remained in year $t$ is divided by the number of sellers per buyer in year $t-1$.
  temp_df1 <- temp_df1 %>%
    transmute(buyer, share = n.y / n.x, buyer_sales) %>%
    summarize(
      mean_share = mean(share),
      mean_share_wt = weighted.mean(share, buyer_sales)
    )

  # the mean is calculated per year weighted and unweighted.

  # then, it's added to the empty dataframe the first time and consecutively per year
  year <- c(i + 1)
  temp_df1 <- cbind(year, temp_df1)
  persistence_buyers <- rbind(persistence_buyers, temp_df1)
}

##
## 2. SELLERS
##


for (i in years) {
  # for each year $t-1$ the set of links between firms is assigned to a dataframe. Then, the number of buyers per seller is calcultated.
  temp_df1 <- sales_new %>%
    filter(year == i) %>%
    select(seller) %>%
    count(seller)

  # the firms of interest are the ones that trade in year $t-1$ and year $t$, we're not interested in the firms that leave the market in year $t$. So, the list of the sellers in year $t-1$ and year $t$ are attached to their respective vectors.
  temp_df2 <- sales_new %>%
    filter(year == i) %>%
    select(seller) %>%
    distinct()

  temp_df3 <- sales_new %>%
    filter(year == i + 1) %>%
    select(seller) %>%
    distinct()

  # then, the sellers that trade in both years are kept
  temp_df2 <- inner_join(temp_df2, temp_df3, by = "seller")

  # the only sellers that are kept from the original list of sellers are the ones which trade in both years, in order to discard the ones that go out of business and measure the persistence between two years.
  temp_df1 <- left_join(temp_df2, temp_df1, by = "seller")

  # another set of the sellers and buyers in the year $t-1$ and the year $t$ is created.
  temp_df4 <- sales_new %>%
    filter(year == i) %>%
    select(buyer, seller)

  temp_df5 <- sales_new %>%
    filter(year == i + 1) %>%
    select(buyer, seller)

  # only the links between firms that are present in the two consecutive years are kept
  temp_df4 <- inner_join(temp_df4, 
                         temp_df5, 
                         by = c("buyer", "seller")) %>%
    distinct() %>%
    count(seller)
  # the number of buyers per seller is computed in the last list

  # the number of buyers per seller that remained trading with eachother in year $t$ is matched with the number of buyers per seller in year $t-1$. The missing values are changed by 0's
  temp_df1 <- left_join(temp_df1, temp_df4, by = "seller")

  temp_df1[is.na(temp_df1)] <- 0

  # the observations are matched and then the share of firms that remained in year $t$ is divided by the number of buyers per seller in year $t-1$.
  temp_df1 <- temp_df1 %>%
    transmute(seller, share = n.y / n.x) %>%
    summarize(mean_share = mean(share))

  # then, it's added to the empty dataframe the first time and consecutively per year
  year <- c(i + 1)
  temp_df1 <- cbind(year, temp_df1)
  persistence_sellers <- rbind(persistence_sellers, temp_df1)
}
```

## Scatter plots with polynomial fit

These histograms correspond to the task 3.2. The first two graphics show the share of productive linkages that survived by buyer between two years, indexed by year. Also, this plot was weighted and unweighted. The second graphic is unweighted and it measures the desired phenomenon by seller. Finally, there's a polynomial fit added to the plots and the degree of them can be manipulated. 

```{r scatter-plots}
degree_poly_buyer <- 1
degree_poly_seller <- 1

##
## 1. BUYERS
##

# unweighted persistence
persist <- ggplot(
  data = persistence_buyers,
  aes(x = year, y = mean_share)
) +
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE,
    formula = y ~ poly(x, degree_poly_buyer)
  ) +
  scale_x_continuous(breaks = seq(2000, 2010, 1)) +
  ggtitle("Unweighted Persistence: Buyers") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(y = "Persistence", x = "Year")
persist

# weighted persistence
persist <- ggplot(
  data = persistence_buyers,
  aes(x = year, y = mean_share_wt)
) +
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE,
    formula = y ~ poly(x, degree_poly_buyer)
  ) +
  scale_x_continuous(breaks = seq(2000, 2010, 1)) +
  ggtitle("Weighted Persistence: Buyers") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(y = "Persistence", x = "Year")
persist

##
## 2. SELLERS
##

# unweighted persistence
persist <- ggplot(
  data = persistence_sellers,
  aes(x = year, y = mean_share)
) +
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE,
    formula = y ~ poly(x, degree_poly_seller)
  ) +
  scale_x_continuous(breaks = seq(2000, 2010, 1)) +
  ggtitle("Unweighted Persistence: Sellers") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(y = "Persistence", x = "Year")
persist
```
