---
title: "Plotting different figures about production linkages"
author: "Jose Ignacio Gonzalez Rojas"
date: "30/05/2020"
output:
  html_document:
    df_print: paged
    number_sections: yes    
    toc: yes
  pdf_document:
    df_print: paged
    number_sections: yes
    toc: yes
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{mathpazo}
---

# General information

The purpose of this code is to create the necessary variables in order to plot a histogram with the count of the sellers per buyer in a specific year, also, this is calculated by sector. These to histograms are also plot with the weights provided by the variable 'buyer_sales'. Also, the share of sellers per buyer and buyers per sellers that he/she still buys/sells from/to between two years. The first one is also weighted by the variable 'buyer_sales'. Finally, the latter variables are plotted in a scatter plot with polynomial fits included. 

## Input files

1. `0-Raw_Data/Sales/sales_new.csv`

## Output files

None. The graphics are kept in code and can be exported to whichever format it's wanted, as well as the dataframes. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
options(tinytex.verbose = TRUE)
```


```{r packages, message=FALSE, warning=FALSE, cache=FALSE, include=FALSE, results='hide'}
# libraries
# vector of libraries to be used

libs <- c(
  "dplyr", "tidyr", "styler", "lintr", "readr", "janitor", "ggplot2",
  "broom", "geosphere", "reshape", "scales", "lubridate", "plotrix"
)

# Installing libraries in case they are not installed
new_packages <- libs[!(libs %in% installed.packages()[, "Package"])]
if (length(new_packages)) install.packages(new_packages)

# Load corresponding libraries

lapply(libs, require, character.only = TRUE)
```

# Importing data to R

In this step, the data needed it's imported into R. Note that the data sets are
in Comma-Separated Values format, so that's why the library "readr" was loaded. The year variable is loaded as DateTime to allow *ggplot2* to detect it as a date. This will make easier the labelling of the x-axis. 

```{r data-importing, echo=TRUE}
# Main inputs

sales_new <- read_csv("0-Raw_Data/Sales/sales_new.csv")
```

## Cleaning the data

The library "janitor" helps standardizing the variable names to the
*snake_case*'s format.

```{r cleaning-variable-names, echo=TRUE}
sales_new <- janitor::clean_names(sales_new)
```

# Number of sellers

## Data processing and statistics

First, the number of sellers per buyer is calculated, this will be helpful to create the plots in adition to the 'buyer_sales' variable as weights. Then, the number of sellers per buyer in each sector is computed. This will be done to a subset of the data corresponding to the year 2001. 

```{r data-manipulation}
yr <- 2001
# note1: better if comments are also in the code
# note2: better if variable names are informative

# number of sellers per buyer
sellers_per_buyer <- sales_new %>%
  filter(year == yr) %>%
  select(buyer, buyer_sales, year) %>%
  group_by(buyer, year) %>%
  mutate(n_sellers = n()) %>%
  unique() %>%
  ungroup()

# number of seller sectors per buyer
sectors_per_buyer <- sales_new %>%
  filter(year == yr) %>%
  select(seller_sector, buyer, buyer_sales, year) %>%
  group_by(buyer, seller_sector, year) %>%
  mutate(n_seller_sectors = n()) %>%
  ungroup() %>%
  select(buyer, buyer_sales, year, n_seller_sectors) %>%
  group_by(buyer, year) %>%
  unique() %>%
  ungroup()

# weighted number of sellers per buyer (?)
# note: each buyer appears as many times as the number
# of sellers already. I think there is double-counting
# in the following formula
sales_new2 <- sales_new %>%
  filter(year == yr) %>%
  select(buyer, seller, buyer_sales) %>%
  count(buyer, wt = buyer_sales)
```

## Histograms

These histograms correspond to the task 3.1. The first two graphics show the frequency of how many sellers per buyer are there, the first is unweighted and the second one is weighted by the variable "buyer_sales". 

```{r histograms}

##
## 1. Number of sellers per buyer
##

## unweighted
# this is another way
h <- hist(sellers_per_buyer$n_sellers,
  main = "Unweighted number of sellers",
  freq = FALSE
)

## weighted
h <- weighted.hist(sellers_per_buyer$n_sellers,
  sellers_per_buyer$buyer_sales,
  freq = FALSE,
  main = "Weighted number of sellers",
  breaks = seq(
    0, max(sellers_per_buyer$n_sellers), max(sellers_per_buyer$n_sellers) / 10
  )
)

graph1 <- ggplot(data = sellers_per_buyer, aes(x = n_sellers)) +
  geom_histogram(bins = 10, boundary = 0) +
  scale_x_continuous(breaks = seq(0, max(sellers_per_buyer$n_sellers), max(sellers_per_buyer$n_sellers) / 10))
graph1

##
## 2. Number of sectors per buyer
##

## unweighted
h <- hist(sectors_per_buyer$n_seller_sectors,
  main = "Unweighted number of sellers",
  breaks = seq(
    0, max(sectors_per_buyer$n_seller_sectors) + 1,
    max(sectors_per_buyer$n_seller_sectors) / 10
  ),
  freq = FALSE
)

## weighted
h <- weighted.hist(
  sectors_per_buyer$n_seller_sectors,
  sectors_per_buyer$buyer_sales,
  freq = FALSE,
  main = "Weighted number of sellers",
  breaks = seq(
    0, max(sectors_per_buyer$n_seller_sectors) + 1,
    max(sectors_per_buyer$n_seller_sectors) / 10
  )
)
```

# persistence of the productive linkages between two year: buyers and sellers. 

## persistance between two years for sellers and buyers. 

This code will work for any time series.  

```{r persistence_productive}
# all the unique years from the 'year' variable are put into a vector
years <- unique(sales_new$year)
# the last year is popped because the for-loop needs to stop in the year $T-1$ to calculate the difference between firms in the year $T$.
years <- years[-length(years)]

# two empty dataframes are created to store the averages of the share of firms that remain from one year to the next one
persistence_buyers <- read.csv(text = "year, mean_share, mean_share_wt")
persistence_sellers <- read.csv(text = "year, mean_share")

##
## 1. BUYERS
##

# variables will be created for every year $t-1$ and the next one $t$ for the buyers and sellers
for (i in years) {
  # for each year $t-1$ the set of links between firms is assigned to a dataframe with their respective weights per buyer corresponding to the buyer_sales in year $t-1$. Then, the number of sellers per buyer is calcultated.
  temp_df1 <- sales_new %>%
    filter(year == i) %>%
    select(buyer, buyer_sales) %>%
    add_count(buyer) %>%
    distinct()

  # the firms of interest are the ones that trade in year $t-1$ and year $t$, we're not interested in the firms that leave the market in year $t$. So, the list of the buyers in year $t-1$ and year $t$ are attached to their respective vectors.
  temp_df2 <- sales_new %>%
    filter(year == i) %>%
    select(buyer) %>%
    distinct()

  temp_df3 <- sales_new %>%
    filter(year == i + 1) %>%
    select(buyer) %>%
    distinct()

  # then, the buyers that trade in both years are kept
  temp_df2 <- inner_join(temp_df2, temp_df3, by = "buyer")

  # from the original list of buyers with their respective weights, the only buyers that are kept are the ones which trade in both years, in order to discard the ones that go out of business. With that in mind, the persistence between two years will be measured.
  temp_df1 <- left_join(temp_df2, temp_df1, by = "buyer")

  # Another set of the sellers and buyers in the year $t-1$ and the year $t$ is created.
  temp_df4 <- sales_new %>%
    filter(year == i) %>%
    select(buyer, seller)

  temp_df5 <- sales_new %>%
    filter(year == i + 1) %>%
    select(buyer, seller)

  # only the links between firms that are present in the two consecutive years are kept
  temp_df4 <- semi_join(temp_df4, temp_df5, by = c("buyer", "seller")) %>%
    count(buyer)
  # the number of sellers per buyer is computed in the last list

  # the number of sellers per buyer that remained trading with eachother in year $t$ is matched with the number of sellers per buyer in year $t-1$. The missing values are changed by 0's
  temp_df1 <- left_join(temp_df1, temp_df4, by = "buyer")

  temp_df1[is.na(temp_df1)] <- 0

  # the observations are matched and then the share of firms that remained in year $t$ is divided by the number of sellers per buyer in year $t-1$.
  temp_df1 <- temp_df1 %>%
    transmute(buyer, share = n.y / n.x, buyer_sales) %>%
    summarize(
      mean_share = mean(share),
      mean_share_wt = weighted.mean(share, buyer_sales)
    )

  # the mean is calculated per year weighted and unweighted.

  # then, it's added to the empty dataframe the first time and consecutively per year
  year <- c(i + 1)
  temp_df1 <- cbind(year, temp_df1)
  persistence_buyers <- rbind(persistence_buyers, temp_df1)
}

##
## 2. SELLERS
##


for (i in years) {
  # for each year $t-1$ the set of links between firms is assigned to a dataframe. Then, the number of buyers per seller is calcultated.
  temp_df1 <- sales_new %>%
    filter(year == i) %>%
    select(seller) %>%
    count(seller)

  # the firms of interest are the ones that trade in year $t-1$ and year $t$, we're not interested in the firms that leave the market in year $t$. So, the list of the sellers in year $t-1$ and year $t$ are attached to their respective vectors.
  temp_df2 <- sales_new %>%
    filter(year == i) %>%
    select(seller) %>%
    distinct()

  temp_df3 <- sales_new %>%
    filter(year == i + 1) %>%
    select(seller) %>%
    distinct()

  # then, the sellers that trade in both years are kept
  temp_df2 <- inner_join(temp_df2, temp_df3, by = "seller")

  # the only sellers that are kept from the original list of sellers are the ones which trade in both years, in order to discard the ones that go out of business and measure the persistence between two years.
  temp_df1 <- left_join(temp_df2, temp_df1, by = "seller")

  # another set of the sellers and buyers in the year $t-1$ and the year $t$ is created.
  temp_df4 <- sales_new %>%
    filter(year == i) %>%
    select(buyer, seller)

  temp_df5 <- sales_new %>%
    filter(year == i + 1) %>%
    select(buyer, seller)

  # only the links between firms that are present in the two consecutive years are kept
  temp_df4 <- semi_join(temp_df4,
    temp_df5,
    by = c("buyer", "seller")
  ) %>%
    count(seller)
  # the number of buyers per seller is computed in the last list

  # the number of buyers per seller that remained trading with eachother in year $t$ is matched with the number of buyers per seller in year $t-1$. The missing values are changed by 0's
  temp_df1 <- left_join(temp_df1, temp_df4, by = "seller")

  temp_df1[is.na(temp_df1)] <- 0

  # the observations are matched and then the share of firms that remained in year $t$ is divided by the number of buyers per seller in year $t-1$.
  temp_df1 <- temp_df1 %>%
    transmute(seller, share = n.y / n.x) %>%
    summarize(mean_share = mean(share))

  # then, it's added to the empty dataframe the first time and consecutively per year
  year <- c(i + 1)
  temp_df1 <- cbind(year, temp_df1)
  persistence_sellers <- rbind(persistence_sellers, temp_df1)
}
```

## Scatter plots with polynomial fit

These histograms correspond to the task 3.2. The first two graphics show the share of productive linkages that survived by buyer between two years, indexed by year. Also, this plot was weighted and unweighted. The second graphic is unweighted and it measures the desired phenomenon by seller. Finally, there's a polynomial fit added to the plots and the degree of them can be manipulated. 

```{r scatter-plots}
poly_degree_buyer <- 1
poly_degree_seller <- 1

##
## 1. BUYERS
##

# unweighted persistence
persist <- ggplot(
  data = persistence_buyers,
  aes(x = year, y = mean_share)
) +
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE,
    formula = y ~ poly(x, poly_degree_buyer)
  ) +
  scale_x_continuous(breaks = seq(2000, 2010, 1)) +
  ggtitle("Unweighted Persistence: Buyers") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(y = "Persistence", x = "Year")
persist

# weighted persistence
persist <- ggplot(
  data = persistence_buyers,
  aes(x = year, y = mean_share_wt)
) +
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE,
    formula = y ~ poly(x, poly_degree_buyer)
  ) +
  scale_x_continuous(breaks = seq(2000, 2010, 1)) +
  ggtitle("Weighted Persistence: Buyers") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(y = "Persistence", x = "Year")
persist

##
## 2. SELLERS
##

# unweighted persistence
persist <- ggplot(
  data = persistence_sellers,
  aes(x = year, y = mean_share)
) +
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE,
    formula = y ~ poly(x, poly_degree_seller)
  ) +
  scale_x_continuous(breaks = seq(2000, 2010, 1)) +
  ggtitle("Unweighted Persistence: Sellers") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(y = "Persistence", x = "Year")
persist
```

# Creation and survival rates of production linkages

## Creation and survival rates

```{r persistence_productive}
# all the unique years from the 'year' variable are put into a vector
years <- unique(sales_new$year)
# the last year is popped because the for-loop needs to stop in the year $T-1$ to calculate the difference between firms in the year $T$.
years <- years[-length(years)]

# two empty dataframes are created to store the averages of the share of firms that remain from one year to the next one
survival_rates <- read.csv(text = "survival_rate, log_buyer_sales")
creation_rates <- read.csv(text = "creation_rate, log_buyer_sales")

##
## 1. survival rates
##

# variables will be created for every year $t-1$ and the next one $t$ for the buyers and sellers
for (i in years) {
  # for each year $t-1$ the set of links between firms is assigned to a dataframe with their respective weights per buyer corresponding to the buyer_sales in year $t-1$. Then, the number of sellers per buyer is calcultated.
  temp_df1 <- sales_new %>%
    filter(year == i) %>%
    select(buyer, buyer_sales) %>%
    add_count(buyer) %>%
    distinct()

  # the firms of interest are the ones that trade in year $t-1$ and year $t$, we're not interested in the firms that leave the market in year $t$. So, the list of the buyers in year $t-1$ and year $t$ are attached to their respective vectors.
  temp_df2 <- sales_new %>%
    filter(year == i) %>%
    select(buyer) %>%
    distinct()

  temp_df3 <- sales_new %>%
    filter(year == i + 1) %>%
    select(buyer) %>%
    distinct()

  # then, the buyers that trade in both years are kept
  temp_df2 <- inner_join(temp_df2, temp_df3, by = "buyer")

  # from the original list of buyers with their respective weights, the only buyers that are kept are the ones which trade in both years, in order to discard the ones that go out of business. With that in mind, the persistence between two years will be measured.
  temp_df1 <- left_join(temp_df2, temp_df1, by = "buyer")

  # Another set of the sellers and buyers in the year $t-1$ and the year $t$ is created.
  temp_df4 <- sales_new %>%
    filter(year == i) %>%
    select(buyer, seller)

  temp_df5 <- sales_new %>%
    filter(year == i + 1) %>%
    select(buyer, seller)

  # only the links between firms that are present in the two consecutive years are kept
  temp_df4 <- inner_join(temp_df4, temp_df5, by = c("buyer", "seller")) %>%
    distinct() %>%
    count(buyer)
  # the number of sellers per buyer is computed in the last list

  # the number of sellers per buyer that remained trading with eachother in year $t$ is matched with the number of sellers per buyer in year $t-1$. The missing values are changed by 0's
  temp_df1 <- left_join(temp_df1, temp_df4, by = "buyer")

  temp_df1[is.na(temp_df1)] <- 0

  # the observations are matched and then the share of firms that remained in year $t$ is divided by the number of sellers per buyer in year $t-1$.
  temp_df1 <- temp_df1 %>%
    transmute(survival_rate = n.y / n.x, log_buyer_sales = log(buyer_sales))

  # then, it's added to the empty dataframe the first time and consecutively per year
  survival_rates <- rbind(survival_rates, temp_df1)
}

##
## 2. creation rates
##

# variables will be created for every year $t-1$ and the next one $t$ for the buyers and sellers
for (i in years) {
  # for each year $t$ the set of buyers is assigned to a dataframe and the number of sellers per buyer is calcultated.
  temp_df1 <- sales_new %>%
    filter(year == i + 1) %>%
    select(buyer) %>%
    count(buyer)

  # in order to compare the creation and survival rates, we need to use weights from the same year, in this case, $t-1$. So we save and to attach them at the end.
  temp_df7 <- sales_new %>%
    filter(year == i) %>%
    select(buyer, buyer_sales) %>%
    distinct(.keep_all = TRUE)

  # the firms of interest are the ones that trade in year $t-1$ and year $t$, we're not interested in the firms that enter the market in year $t$. So, the list of the buyers in year $t-1$ and year $t$ are attached to their respective vectors.
  temp_df2 <- sales_new %>%
    filter(year == i) %>%
    select(buyer) %>%
    distinct()

  temp_df3 <- sales_new %>%
    filter(year == i + 1) %>%
    select(buyer) %>%
    distinct()

  # then, the buyers that trade in both years are kept
  temp_df2 <- inner_join(temp_df2, temp_df3, by = "buyer")

  # from the original list of buyers with their respective weights, the only buyers that are kept are the ones which trade in both years, in order to discard the ones that go out of business. With that in mind, the persistence between two years will be measured.
  temp_df1 <- left_join(temp_df2, temp_df1, by = "buyer")

  # Another set of the sellers and buyers in the year $t-1$ and the year $t$ is created.
  temp_df4 <- sales_new %>%
    filter(year == i) %>%
    select(buyer, seller)

  temp_df5 <- sales_new %>%
    filter(year == i + 1) %>%
    select(buyer, seller)

  # only the links between firms that are present in year $t$ and aren't in year $t-1$ are kept
  temp_df4 <- anti_join(temp_df5, temp_df4, by = c("buyer", "seller")) %>%
    count(buyer)
  # the number of sellers per buyer is computed in the last list

  # the number of sellers per buyer that started trading with eachother in year $t$ is matched with the number of sellers per buyer in year $t-1$. The missing values are changed by 0's
  temp_df1 <- left_join(temp_df1, temp_df4, by = "buyer")

  temp_df1 <- left_join(temp_df1, temp_df7, by = "buyer")

  temp_df1[is.na(temp_df1)] <- 0

  # the observations are matched and then the share of firms that remained in year $t$ is divided by the number of sellers per buyer in year $t-1$.
  temp_df1 <- temp_df1 %>%
    transmute(creation_rate = n.y / n.x, log_buyer_sales = log(buyer_sales))

  # then, it's added to the empty dataframe the first time and consecutively per year
  creation_rates <- rbind(creation_rates, temp_df1)
}
```

## Scatter plots for survival and creation rates

```{r scatter-plots}
poly_degree_survival <- 1
poly_degree_creation <- 1

##
## 1. SURVIVAL RATE
##

rates <- ggplot(
  data = survival_rates,
  aes(x = log_buyer_sales, y = survival_rate)
) +
  geom_point(size = 0.001) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    formula = y ~ poly(x, poly_degree_survival)
  ) +
  ggtitle("Survival rate of production linkages") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(y = "Persistence rate", x = "Log of buyer sales")
rates


##
## 2. CREATION RATE
##

rates <- ggplot(
  data = creation_rates,
  aes(x = log_buyer_sales, y = creation_rate)
) +
  geom_point(size = 0.001) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    formula = y ~ poly(x, poly_degree_creation)
  ) +
  ggtitle("Creation rate of production linkages") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(y = "Creation rate", x = "Log of buyer sales")
rates
```

# Survival function

## Approximating the survival function
```{r survival-function}
# write hashes in front of this code when working with the real data, this is just a fix for the fake dataset

survival_function <- sales_new %>%
  group_by(buyer, seller, year) %>%
  distinct(buyer, seller, year, .keep_all = TRUE)

diff_tolerance <- 1 # this parameter indicates how long can two firms stay without trading and don't break their link

survival_function <- survival_function %>%
  select(year, buyer, seller) %>% # selecting the variables of interest
  group_by(buyer, seller) %>% # separating the links between firms
  arrange(year) %>% # sorting ascendently by year
  mutate(diff = year - lag(year)) %>% # calculating the first order difference between the year variable
  mutate(diff = if_else(diff > diff_tolerance + 1, 0, diff)) # if the difference is larger than the tolerance plus one, we create a new link

survival_function[is.na(survival_function)] <- 0 # the NA values correspond to the starting year of the first link between two firms, so it's set as 0 to indicate it's a starting year for the corresponding link

survival_function <- survival_function %>%
  group_by(buyer, seller) %>% # separating the links between firms
  mutate(link_number = cumsum(c(-1, diff(diff)) < 0)) %>% # if there's intermittent trading between two firms,the new links created, respecting the tolerance, are separated
  ungroup() %>%
  group_by(buyer, seller, link_number) %>% # regroup taking into account the new separation of the links between firms
  arrange(year) %>% # sorting ascendently again by year
  mutate(yr = seq(n()) - 1) %>% # assigning the year from 0 to 4 according to the length of the firms relationship
  filter(!any(yr == 0 & year > 2005) & yr <= 4) %>% # censoring the data: if the observation is greater or equal than year 4 it's not taken into account; the starting links after 2005 are not considered in this exercise
  ungroup() %>%
  select(yr) %>%
  count(yr) # counting the links that remain each year

yr_0_total_links <- survival_function$n[1] # saving the base for the survival rate

survival_function <- survival_function %>%
  transmute(year = yr, survival_rate = n / yr_0_total_links) # resulting df with year from 0 to 4 and their respecting share
```

## Plotting the survival function
```{r plotting-survival-function}
surv_func <- ggplot(
  data = survival_function,
  aes(x = year, y = survival_rate)
) +
  geom_bar(stat = "identity") +
  ggtitle("Survival function of the links between firms \n Survival rate after year t") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(y = "Survival rate", x = "Year")
surv_func
```

## Weighted survival function: three possible options of weighting

```{r weighted-survival-function}
# write hashes in front of this code when working with the real data, this is just a fix for the fake dataset

weighted_survival_function <- sales_new %>%
  group_by(buyer, seller, year) %>%
  distinct(buyer, seller, year, .keep_all = TRUE)

diff_tolerance <- 1 # this parameter indicates how long can two firms stay without trading and don't break their link

weighted_survival_function <- weighted_survival_function %>%
  select(year, buyer, seller, buyer_sales) %>% # selecting the variables of interest
  group_by(buyer, seller) %>% # separating the links between firms
  arrange(year) %>% # sorting ascendently by year
  mutate(diff = year - lag(year)) %>% # calculating the first order difference between the year variable
  mutate(diff = if_else(diff > diff_tolerance + 1, 0, diff)) # if the difference is larger than the tolerance plus one, we create a new link

weighted_survival_function[is.na(weighted_survival_function)] <- 0 # the NA values correspond to the starting year of the first link between two firms, so it's set as 0 to indicate it's a starting year for the corresponding link

weighted_survival_function <- weighted_survival_function %>%
  group_by(buyer, seller) %>% # separating the links between firms
  mutate(link_number = cumsum(c(-1, diff(diff)) < 0)) %>% # if there's intermittent trading between two firms,the new links created, respecting the tolerance, are separated
  ungroup() %>%
  group_by(buyer, seller, link_number) %>% # regroup taking into account the new separation of the links between firms
  arrange(year) %>% # sorting ascendently again by year
  mutate(yr = seq(n()) - 1) %>% # assigning the year from 0 to 4 according to the length of the firms relationship
  filter(!any(yr == 0 & year > 2005) & yr <= 4) %>% # censoring the data: if the observation is greater or equal than year 4 it's not taken into account; the starting links after 2005 are not considered in this exercise
  ungroup()

# put hashes in front of the options that aren't picked

# Option 1: leave the weights as they are

weighted_survival_function <- weighted_survival_function %>%
  select(-year, -diff)

# Option 2: apply the buyer_sales variable from year 0 to the other ones

# first, the buyer_sales variable from year 0 is written in a dataframe with their respective buyer, seller and link_number

temp_df1 <- weighted_survival_function %>%
  select(-year, -diff) %>%
  filter(yr == 0) %>%
  select(yr, buyer, seller, link_number, buyer_sales) %>%
  distinct(.keep_all = TRUE)

# these weights are matched with their respective links for years $t$, with t in 1, 2, 3, 4

temp_df2 <- weighted_survival_function %>%
  filter(yr != 0) %>%
  select(-year, -diff) %>%
  left_join(temp_df1, by = c("buyer", "seller", "link_number")) %>% # for each link the respective weight is assigned
  transmute(yr = yr.x, buyer, seller, link_number, buyer_sales = buyer_sales.y) # given we have both values for buyer_sales and yr at the same time, the desired ones are kept

weighted_survival_function <- rbind(temp_df1, temp_df2) # appending the year 0 with the rest

# Option 3: apply the buyer_sales variable from year $t-1$ to year $t$ for $t$ >= 1

temp_df1 <- weighted_survival_function %>%
  filter(yr == 0) %>%
  select(-year, -diff) # selecting the variables of interest

for (i in 1:4) {
  # a dataframe with the weights of year $t - 1$ is created
  temp_df2 <- weighted_survival_function %>%
    filter(yr == i - 1) %>%
    select(buyer, seller, link_number, buyer_sales) %>%
    distinct(.keep_all = TRUE)

  # these weights are matched to year $t$
  temp_df3 <- weighted_survival_function %>%
    filter(yr == i) %>%
    select(-year, -diff) %>%
    left_join(temp_df2, by = c("buyer", "seller", "link_number")) %>%
    transmute(yr, buyer, seller, link_number, buyer_sales = buyer_sales.y) # given we have both values for buyer_sales at the same time, the desired ones are kept

  temp_df1 <- rbind(temp_df1, temp_df3) # the links with their correct weights are appended to the dataframe
}

weighted_survival_function <- temp_df1

#
# Summarizing and creating the weighted survival function
#
```

## Plotting the weighted survival function
```{r plotting-weighted-survival-function}
# The next graph needs a summarized dataframe with the year with their respective survival rates, if that's not the case feel free to modify it

surv_func <- ggplot(
  data = survival_function,
  aes(x = year, y = survival_rate)
) +
  geom_bar(stat = "identity") +
  ggtitle("Survival function of the links between firms weighted by buyer sales \n Survival rate after year t") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(y = "Survival rate", x = "Year")
surv_func
```

## Checking if the survival rate from task 3.4 if $t = 1$ is the same as the persistance rate from one year to another in task 3.3

```{r checking-equality-of-rates}
# persistence rate from task 3.3

mean(survival_rates$survival_rate)

# survival function S(x) with $t = 1$ S(1)

survival_function$survival_rate[2]
```
