---
title: "Plotting a figure of shares of sales in different provinces"
author: "José Ignacio González Rojas"
date: "19/5/2020"
output: 
  pdf_document:
    toc: yes
    df_print: paged
    number_sections: yes
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{mathpazo}
---

# General information

The purpose of this code is divided in two main objectives. 

The first one is to calculate the number of sellers from province P to province P'. Then, to compute the share of the largest seller of province P to province P'. Also, the share of sellers in PP' that have transactions among each other. Aditionally, the distances in between the provinces P and P'. As well as the share of the largest seller of the sector S to sector S'. Finally, multiple regressions by the pair of sectors SS' to explain the decision of a seller firm in a specific sector to trade with a buyer firm in the sector S' explained by the log of the distance between the provinces P and P'. 

The second objective is to plot an area graphic with the provinces in the X-axis, sorted ascendently by the share of the largest seller of the province P to the province P', and in the Y-axis the number of sellers in the provinces. In adition to plot the constants of the regressions with respect to the share of the largest seller in SS'. 

## Input files

1. `0-Raw_Data/Sales/sales_states.csv` The file has 6 columns. The year, the identity of the firm that sells the product "seller", the identity of the firm that buys the product "buyer", the province where the seller lives "state_seller", the province where the buyer lives "state_buyer", and finally, the amount of the sale "amount". 
2. `0-Raw_Data/Sales/coordinates.csv` The file has 3 columns: the state, the longitude and latitude. 

## Output files

1. `3-Graphics/a1.eps` This file contains a graphic with the provinces in the X-axis, sorted ascendently by the share of the largest seller of the province P to the province P', and in the Y-axis the number of sellers in the provinces. 
2. `3-Graphics/a1.png`
3. `3-Graphics/a1.pdf`
4. `3-Graphics/a2.eps`
5. `3-Graphics/a2.png`
6. `3-Graphics/a2.pdf`
7. `3-Graphics/b.eps`
8. `3-Graphics/b.png`
9. `3-Graphics/b.pdf`
10. `3-Graphics/l1.eps`
11. `3-Graphics/l1.png`
12. `3-Graphics/l1.pdf`
13. `3-Graphics/l2.eps`
14. `3-Graphics/l2.png`
15. `3-Graphics/l2.pdf`
16. `3-Graphics/a1_s.eps` This file contains a graphic with the sector in the X-axis, sorted ascendently by the share of the largest seller of the sector P to the sector P', and in the Y-axis the constant of the respective regression.
17. `3-Graphics/a1_s.png`
18. `3-Graphics/a1_s.pdf`
19. `3-Graphics/a2_s.eps`
20. `3-Graphics/a2_s.png`
21. `3-Graphics/a2_s.pdf`
22. `3-Graphics/b_s.eps`
23. `3-Graphics/b_s.png`
24. `3-Graphics/b_s.pdf`
25. `3-Graphics/l1_s.eps`
26. `3-Graphics/l1_s.png`
27. `3-Graphics/l1_s.pdf`
28. `3-Graphics/l2_s.eps`
29. `3-Graphics/l2_s.png`
30. `3-Graphics/l2_s.pdf`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
options(tinytex.verbose = TRUE)
```


```{r packages, message=FALSE, warning=FALSE, cache=FALSE, include=FALSE, results='hide'}
# libraries
# vector of libraries to be used

libs <- c("dplyr", "tidyr", "styler", "lintr", "readr", "janitor", "ggplot2", "lme4", "geosphere", "reshape")

# Installing libraries in case they are not installed

for(i in length(libs)){ 
  if(!(libs[i] %in% installed.packages())) install.packages(libs) 
}

# Load corresponding libraries

lapply(libs, require, character.only=TRUE)
```

# Importing data to R

In this step, the data needed it's imported into R. Note that the data sets are in Comma-Separated Values format, so that's why the library "readr" was loaded.

```{r data-importing, echo=TRUE}
# Main inputs

sales_states <- read_csv("0-Raw_Data/Sales/sales_states.csv")
coordinates1 <- read_csv("0-Raw_Data/Sales/coordinates.csv")
```

# Cleaning the data

## Cleaning variable names

The library "janitor" helps standardizing the variable names to the *snake_case*'s format. 

```{r cleaning-variable-names, echo=TRUE}
sales_states <- janitor::clean_names(sales_states)
coordinates1 <- janitor::clean_names(coordinates1)
```

# Manipulating the data

Before plotting the requested graphics, the data needs to manipulated in order to create some variables. 

## Computing the number of sellers in the observation PP'

The number of sellers from province P to province P' is calculated by counting the seller that made at least one transaction. 

```{r number-of-sellers, echo=TRUE}
number_of_sellers <- sales_states %>%
  group_by(state_seller, state_buyer) %>%
  count(seller) %>% 
  count() %>%
  ungroup()
```

## Calculating the share of the largest seller in the observation PP'

The share of the largest seller in the observation PP' calculus can be splitted in two parts. First, the sum of all the sales of a seller i in the observation PP'. Then, the sales of everyone in the observation PP'. Finally, the shares are calculated and the maximum observation is selected. 

```{r share-of-the-largest-seller, echo=TRUE}
share_largest_seller <- sales_states %>%
  group_by(state_seller, state_buyer, seller) %>% 
  mutate(seller_sales = sum(amount)) %>% 
  ungroup() %>% 
  group_by(state_seller, state_buyer) %>%
  transmute(share = seller_sales / sum(amount)) %>% 
  select(state_seller, state_buyer, share) %>% 
  top_n(1, share) %>% 
  unique()
```

## Computing the share of sellers in the observation PP'

The share of the sellers in the observation PP' is the number of sellers divided by the number of observations PP'. 

```{r share-of-the-largest-seller, echo=TRUE}
share_of_sellers <- number_of_sellers %>% 
  mutate(total = n()) %>% 
  mutate(share_sellers = n / total)
```

## Expanding the data set

Here the data set of the coordinates it's expanded and all the possible combinations are found. First, an identical data set it's needed with different column names. 

```{r expansion, echo=TRUE}
# Duplicate the dataset

coordinates2 <- coordinates1

colnames(coordinates1) <- paste(colnames(coordinates1), "1", sep="") #Adds 1 at the end of the variable name
colnames(coordinates2) <- paste(colnames(coordinates2), "2", sep="") #Adds 2 at the end of the variable name

coordinates <- expand.grid.df(coordinates1, coordinates2) # Combines all the rows in every possible way 
```

## Calculating the distances in the observations PP'

The distances between two provinces P and P' are calculated with the standard formula. If the distance between two provinces is zero, it's changed to 1. 

```{r share-of-the-largest-seller, echo=TRUE}
distances <- coordinates %>% 
  rowwise() %>% # The next operation can't be made with matrix vectorization, that's why it's computed row by row. 
  mutate(dist = distGeo(c(lon1, lat1), c(lon2, lat2)) / 1000) %>% 
  mutate(dist = if_else(dist == 0, 1, dist))%>% 
  mutate(dist = log(dist))%>% 
  ungroup() %>% #The reverse command of rowwise()
  select(-lon1, -lon2, -lat1, -lat2)
```

## Joining the data sets

The datasets are joined by the "state_seller" and "state_buyer" variables and the "n" and "share" variables are kept in the first case and the "share_of_sellers" and "share" in the second one. Also, a the data is sorted to comply with the plot structure. This is accomplished by creating the variable "xpp" which is a sequence from 1 to 400.

```{r joining-the-datasets, echo=TRUE}
final <- merge(number_of_sellers, share_largest_seller) %>% 
  arrange(share) %>% 
  mutate(xpp = seq(n())) %>% 
  rename(npp = n)

final_s <- merge(share_of_sellers, share_largest_seller) %>% 
  arrange(share) %>% 
  mutate(xpp = seq(n())) %>% 
  rename(npp = share_sellers)
```

# Plotting the data

First, three options of graphics are given: area, line and bar plots. The line and area plots have different versions. Then, the graphics are exported to .eps and .png files. 

## Bar plot 

```{r bar-plot, echo=TRUE}
base <- ggplot(data = final, aes(x = xpp, y = npp)) 
b <- base + geom_bar(stat = "identity", fill="#0078b5", width=0.05)
b

base_s <- ggplot(data = final_s, aes(x = xpp, y = npp)) 
b_s <- base_s + geom_bar(stat = "identity", fill="#0078b5", width=0.05)
b_s
```

## Line plot 

```{r line-plot, echo=TRUE}
l1 <- base + geom_line(color="#0078b5", size = 0.25) + ylim(0, 61)
l1
l1_s <- base_s + geom_line(color="#0078b5", size = 0.25) + ylim(0, 1)
l1_s
l2 <- base + geom_line(color="#0078b5", size = 0.25)
l2
l2_s <- base_s + geom_line(color="#0078b5", size = 0.25)
l2_s
```
## Area plot 

```{r area-plot, echo=TRUE}
a1 <- base + geom_area(fill="#0078b5", alpha=0.75)
a1
a1_s <- base_s + geom_area(fill="#0078b5", alpha=0.75)
a1_s
a2 <- base + geom_area(fill="#0078b5", alpha=0.75) + geom_line(color="#0078b5", size = 0.25)
a2
a2_s <- base_s + geom_area(fill="#0078b5", alpha=0.75) + geom_line(color="#0078b5", size = 0.25)
a2_s
```

## Exporting the graphics

Each of the graphics are exported to .eps and .png files. 

```{r exporting-the-graphics, echo=TRUE}
formats <- c("eps", "png", "pdf")
for(i in formats){
  ggsave(filename = paste0("a1.",i), plot = a1, path = "3-Graphics", device = i, width = 7.55, height = 4.25, dpi = 1200)
  ggsave(filename = paste0("a2.",i), plot = a2, path = "3-Graphics", device = i, width = 7.55, height = 4.25, dpi = 1200)
  ggsave(filename = paste0("b.",i), plot = b, path = "3-Graphics", device = i, width = 7.55, height = 4.25, dpi = 1200)
  ggsave(filename = paste0("l1.",i), plot = l1, path = "3-Graphics", device = i, width = 7.55, height = 4.25, dpi = 1200)
  ggsave(filename = paste0("l2.",i), plot = l2, path = "3-Graphics", device = i, width = 7.55, height = 4.25, dpi = 1200)
  ggsave(filename = paste0("a1_s.",i), plot = a1_s, path = "3-Graphics", device = i, width = 7.55, height = 4.25, dpi = 1200)
  ggsave(filename = paste0("a2_s.",i), plot = a2_s, path = "3-Graphics", device = i, width = 7.55, height = 4.25, dpi = 1200)
  ggsave(filename = paste0("b_s.",i), plot = b_s, path = "3-Graphics", device = i, width = 7.55, height = 4.25, dpi = 1200)
  ggsave(filename = paste0("l1_s.",i), plot = l1_s, path = "3-Graphics", device = i, width = 7.55, height = 4.25, dpi = 1200)
  ggsave(filename = paste0("l2_s.",i), plot = l2_s, path = "3-Graphics", device = i, width = 7.55, height = 4.25, dpi = 1200)
}
```